---
title: "Getting started with nonabsdid"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting started with nonabsdid}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7,
  fig.height = 4.5
)
```

```{r setup}
library(nonabsdid)
```

`nonabsdid` provides three things on top of the heterogeneity-robust DiD
estimators that already exist in R:

1. A unified front door, `nabs_event_study()`, that runs DCDH, PanelMatch, or
   any of the `fect` family (FE / IFE / MC) with a common argument set.
2. An S3 generic, `as_nabs_event_study()`, that converts each estimator's
   native output into a tidy tibble with a stable schema.
3. A plotting function, `nabs_event_plot()`, that overlays any combination
   of those tidy objects on a single ggplot2 panel, with an optional naive
   TWFE reference series.

Plus a one-line shortcut, `nabs_event_study_simple()`, that runs all three
heterogeneity-robust estimators with sensible defaults and gives you a
single overlay plot — meant for the very first look at the data.

The estimator packages themselves (`DIDmultiplegtDYN`, `PanelMatch`, `fect`,
`fixest`) are **suggested**, not required, so install only the ones you plan
to use.

## The 30-second version

```{r quick, eval = FALSE}
res <- nabs_event_study_simple(sim,
                               outcome = "y", treatment = "d",
                               unit = "id", time = "t")
res$plot
```

That's it. Window lengths are auto-picked from the panel. All available
heterogeneity-robust estimators run (skipped silently if their package
isn't installed), the naive TWFE reference is overlaid in grey, and you
get a single figure to inspect.

For everything below — picking specific estimators, wider control over
options, comparing FE vs IFE vs MC — read on.

## A simulated panel

We simulate a panel with non-absorbing treatment so that all three estimators
have something to chew on. Half of the units turn on at `t = 10` and half of
those turn back off at `t = 16`.

```{r sim, eval = FALSE}
set.seed(2026)
N <- 80; TT <- 24
sim <- expand.grid(id = seq_len(N), t = seq_len(TT))

# Treatment switches on at t=10 for ids <= N/2, and off at t=16 for ids <= N/4.
sim$d <- with(sim, as.integer(
  (id <= N/2 & t >= 10 & !(id <= N/4 & t >= 16))
))

# Heterogeneous, time-varying effect: 1 for early, 0.5 for late.
sim$tau <- with(sim, ifelse(id <= N/4, 1.0, 0.5))
sim$y <- with(sim, 0.05 * id + 0.03 * t + d * tau + rnorm(nrow(sim), sd = 0.3))
```

## Three estimators, three lines of code

```{r run, eval = FALSE}
res_dcdh <- nabs_event_study(sim, outcome = "y", treatment = "d",
                         unit = "id", time = "t",
                         method = "DCDH", lags = 4, leads = 6)

res_pm   <- nabs_event_study(sim, outcome = "y", treatment = "d",
                         unit = "id", time = "t",
                         method = "PanelMatch", lags = 4, leads = 6)

res_ife  <- nabs_event_study(sim, outcome = "y", treatment = "d",
                         unit = "id", time = "t",
                         method = "IFE")
```

Each `nabs_event_study()` return is a list with `tidy` (an `nabs_event_study_tbl`),
`fit` (the native estimator object, for diagnostics), and `call`.

## Or call the estimators directly

If you want full control of estimator-specific options, call the underlying
package and tidy the result:

```{r direct, eval = FALSE}
fit <- DIDmultiplegtDYN::did_multiplegt_dyn(
  df = sim, outcome = "y", group = "id", time = "t",
  treatment = "d", effects = 6, placebo = 4,
  cluster = "id"
)
tidy_dcdh <- as_nabs_event_study(fit, outcome = "y")
```

For `PanelMatch`, remember to pass the placebo result via `pre_obj`:

```{r pm-direct, eval = FALSE}
panel <- PanelMatch::PanelData(sim, "id", "t", "d", "y")
pm    <- PanelMatch::PanelMatch(panel.data = panel, lag = 4, lead = 0:6,
                                refinement.method = "ps.match",
                                size.match = 10, qoi = "att",
                                placebo.test = TRUE,
                                forbid.treatment.reversal = FALSE)
pe    <- PanelMatch::PanelEstimate(pm, panel.data = panel)
pl    <- PanelMatch::placebo_test(pm, panel.data = panel, plot = FALSE)

tidy_pm <- as_nabs_event_study(pe, pre_obj = pl)
```

## Naive TWFE reference

For a visual baseline, fit a naive TWFE event study with `fixest`:

```{r twfe, eval = FALSE}
ref <- naive_twfe(sim, outcome = "y", treatment = "d",
                  unit = "id", time = "t",
                  lags = 4, leads = 6)
```

The reference is built from treatment *changes* (a distributed-lag
specification), so it handles the on/off treatment simulated above rather than
assuming a single absorbing onset. It is still intended only as a reference:
it is not robust to treatment-effect heterogeneity -- which is the point.
Showing it next to the robust estimators makes the correction visible.

## Plot

```{r plot, eval = FALSE}
nabs_event_plot(
  res_dcdh$tidy, res_pm$tidy, res_ife$tidy,
  reference = ref,
  xlim = c(-4, 6),
  ylim = c(-1, 2),
  ylab = "Effect on y"
)
```

The reference series is drawn in `grey20` (configurable via `reference_color`)
and dashed, so it sits visually behind the main estimators. Pre-treatment
periods get round markers; post-treatment get triangles. Each method gets its
own color pair; the default palette is patterned after the conventions in
applied DCDH/PanelMatch papers, with red shades pre and blue/green post.

## Schema

Every tidier returns a tibble with the same columns:

| column      | type | description                                             |
|-------------|------|---------------------------------------------------------|
| `time`      | int  | Relative period (0 = treatment onset).                  |
| `estimate`  | num  | Point estimate.                                         |
| `std.error` | num  | Standard error (may be `NA`).                           |
| `conf.low`  | num  | Lower CI bound.                                         |
| `conf.high` | num  | Upper CI bound.                                         |
| `window`    | chr  | `"pre"` if `time < 0`, else `"post"`.                   |
| `method`    | chr  | `"DCDH"`, `"PanelMatch"`, `"IFE"`, `"TWFE"`, or custom. |
| `outcome`   | chr  | Outcome variable name.                                  |

Anything you can coerce to a data frame with at least `time` and `estimate`
columns can also flow through `as_nabs_event_study()`. This makes it easy to plug
in additional estimators later -- write a one-line method that pulls the
right slots, and the plotting code keeps working.
